Title Goes Here

### 2020-02-26 paper/paper.tex
Fixed typo in rewrite rule in paper
Fixed another typo in paper rewrite transformation
Modified wording (also, formatting)

Despite these advantages Haskell provides for creating EDSLs, there are a few
constructs for which representations have proven illusive. One prominent example
is that of a \ttt{case} expression. Pattern matching is a convenient way to
implement control flow structures and to view components of data structures, so
it is frequently a desirable feature for many EDSLs.

@@

In this EDSL, the type representing values in the expression language is \ttt{E t}
and there are two basic functions which translate values to and from this
expression language with the following type signatures:

@@

[\ldots]

@@

\begin{align*}
  T'_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ case s of \{ ...$_A$ \}; ...$_B$ \}}}\\
    &\rewrites
        \ttt{\{ K x0 ... xN $\rarr$ $T_{\ttt{f}}\expr{\ttt{case s of \{ ...$_A$ \}}}$; $T'_{\ttt{f}}\expr{\ttt{\{ ...$_B$ \}}}$\}}\\
  T'_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}}\\
    &\rewrites
      \begin{cases}
        \ttt{\{ K x0 ... xN $\rarr$ body}[\ttt{f} \mapsto \ttt{Step}]\ttt{; }T'_{\ttt{f}}\expr{\ttt{\{ ... \}}}\ttt{ \}} & \text{if \ttt{f} occurs free in \ttt{body}}\\
        \ttt{\{ K x0 ... xN $\rarr$ Done body; ... \}} & \text{otherwise}
      \end{cases}\\
  T'_{\ttt{f}}&\expr{\ttt{\{ \}}} \rewrites \ttt{\{ \}}\\
  T_{\ttt{f}}&\expr{\ttt{case s of \{ ... \}}}\\
    &\rewrites
      \begin{cases}
        \ttt{case s of $T'_{\ttt{f}}\expr{\ttt{\{ ... \}}}$} & \text{if \ttt{f} occurs free in \ttt{\{ ... \}}}\\
        \ttt{Done (case s of \{ ... \})} & \text{otherwise}
      \end{cases}
\end{align*}
\begin{align*}
  &C^{*}\expr{\ttt{x}} \rewrites
    \begin{cases}
      \ttt{x} & \text{if \ttt{x} has no subexpressions marked with \ttt{externalize}}\\
      C^{*}\expr{C\expr{\ttt{y}}} & \text{if \ttt{externalize y} is the first marked subexpression of \ttt{x}}\\
    \end{cases}\\
\end{align*}
\begin{align*}
  &C\expr{\ttt{runIter x}} \rewrites \ttt{TailRec $M\expr{\ttt{x}}$}\\
  &C\expr{\ttt{x + y}} \rewrites \ttt{Add $M\expr{\ttt{x}}$ $M\expr{\ttt{y}}$}\\
  &C\expr{\ttt{case scrutinee of \{ ... \}}} \rewrites \ttt{CaseExp $M\expr{\ttt{scrutinee}}$ } B\expr{\ttt{\{ ... \}}}\\
  &C\expr{\ttt{($\lambda$x $\rarr$ body)}} \rewrites A\expr{\lambda \ttt{x} \rarr \ttt{body}}\\
  &C\expr{\ttt{(f x)}} \rewrites \ttt{ConstructAp $M\expr{\ttt{f}}$ $M\expr{\ttt{x}}$}\;\;\;\;\text{(Where \ttt{f} is not a primitive or an application of a primitive)}\\
  &C\expr{\ttt{x}} \rewrites \ttt{Construct x}\;\;\;\;\text{(Where \ttt{x} is a variable which is not a primitive)}\\
  &\\
  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}} \rewrites \ttt{SumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body0}}\; B\expr{\ttt{\{ ... \}}}\\
  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body \}}} \rewrites \ttt{OneSumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body}}\\
  &B\expr{\ttt{\{ \}}} \rewrites \ttt{EmptyMatch}\\
  &\\
  &P\expr{\ttt{x0 x1 ... $\rarr$ body}} \rewrites \ttt{ProdMatchExp }A\expr{\ttt{$\lambda$ x0 $\rarr$ $P\expr{\ttt{x1 ... $\rarr$ body}}$}}\\
  &P\expr{\ttt{x $\rarr$ body}} \rewrites \ttt{OneProdMatchExp }A\expr{\ttt{$\lambda$ x $\rarr$ body}}\\
  &P\expr{\ttt{ $\rarr$ body}} \rewrites \ttt{NullaryMatch $M\expr{\ttt{body}}$}\\
  &\\
  &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} \rewrites \ttt{Lam (Name @a uniq) $M\expr{\ttt{body$[\ttt{x} \mapsto \ttt{Var uniq}]$}}$}\text{\;\;\;\;\;(where \ttt{uniq} is a globally unique identifier)}\\
  &\\
  &M\expr{\ttt{(x :: a)}} \rewrites
    \begin{cases}
      \ttt{externalize x} & \text{if}\; \nexists t, a\; \typeeq E\; t\\
      \ttt{x} & \text{if}\; \exists t, a\; \typeeq E\; t
    \end{cases}
\end{align*}

>>>>

Despite these advantages Haskell provides for creating EDSLs, there are a few
constructs for which representations have proven illusive. One prominent example
is that of a \ttt{case} expression. Pattern matching is a convenient way to
implement control flow structures and to inspect data structures, so it is
frequently a desirable feature for many EDSLs.

@@

In this EDSL, the type representing values in the expression language is \ttt{E
t} and there are two basic functions which translate values to and from this
expression language with the following type signatures (omitting type class
constraints for the moment):

@@

@@

\begin{align*}
  T'_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ case s of \{ ...$_A$ \}; ...$_B$ \}}}\\
    &\rewrites
        \ttt{\{ K x0 ... xN $\rarr$ $T_{\ttt{f}}\expr{\ttt{case s of \{ ...$_A$ \}}}$; $T'_{\ttt{f}}\expr{\ttt{\{ ...$_B$ \}}}$\}}\\
  T'_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}}\\
    &\rewrites
      \begin{cases}
        \ttt{\{ K x0 ... xN $\rarr$ body0}[\ttt{f} \mapsto \ttt{Step}]\ttt{; }T'_{\ttt{f}}\expr{\ttt{\{ ... \}}}\ttt{ \}} & \text{if \ttt{f} occurs free in \ttt{body0}}\\
        \ttt{\{ K x0 ... xN $\rarr$ Done body0; $T'_{\ttt{f}}\expr{\ttt{\{ ... \}}} \}$} & \text{otherwise}
      \end{cases}\\
  T'_{\ttt{f}}&\expr{\ttt{\{ \}}} \rewrites \ttt{\{ \}}\\
  T_{\ttt{f}}&\expr{\ttt{case s of \{ ... \}}}\\
    &\rewrites
      \begin{cases}
        \ttt{case s of $T'_{\ttt{f}}\expr{\ttt{\{ ... \}}}$} & \text{if \ttt{f} occurs free in \ttt{\{ ... \}}}\\
        \ttt{Done (case s of \{ ... \})} & \text{otherwise}
      \end{cases}
\end{align*}
\begin{align*}
  &C^{*}\expr{\ttt{x}} \rewrites
    \begin{cases}
      \ttt{x} & \text{if \ttt{x} has no subexpressions marked with \ttt{externalize}}\\
      C^{*}\expr{C\expr{\ttt{y}}} & \text{if \ttt{externalize y} is the first marked subexpression of \ttt{x}}\\
    \end{cases}\\
\end{align*}
\begin{align*}
  &C\expr{\ttt{runIter x}} \rewrites \ttt{TailRec $M\expr{\ttt{x}}$}\\
  &C\expr{\ttt{x + y}} \rewrites \ttt{Add $M\expr{\ttt{x}}$ $M\expr{\ttt{y}}$}\\
  &C\expr{\ttt{case scrutinee of \{ ... \}}} \rewrites \ttt{CaseExp $M\expr{\ttt{scrutinee}}$ } B\expr{\ttt{\{ ... \}}}\\
  &C\expr{\ttt{($\lambda$x $\rarr$ body)}} \rewrites A\expr{\lambda \ttt{x} \rarr \ttt{body}}\\
  &C\expr{\ttt{(f x)}} \rewrites \ttt{ConstructAp $M\expr{\ttt{f}}$ $M\expr{\ttt{x}}$}\;\;\;\;\text{(Where \ttt{f} is not a primitive or an application of a primitive)}\\
  &C\expr{\ttt{x}} \rewrites \ttt{Construct x}\;\;\;\;\text{(Where \ttt{x} is a variable which is not a primitive)}\\
  &\\
  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}} \rewrites \ttt{SumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body0}}\; B\expr{\ttt{\{ ... \}}}\\
  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body \}}} \rewrites \ttt{OneSumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body}}\\
  &B\expr{\ttt{\{ \}}} \rewrites \ttt{EmptyMatch}\\
  &\\
  &P\expr{\ttt{x0 x1 ... $\rarr$ body}} \rewrites \ttt{ProdMatchExp }A\expr{\ttt{$\lambda$ x0 $\rarr$ $P\expr{\ttt{x1 ... $\rarr$ body}}$}}\\
  &P\expr{\ttt{x $\rarr$ body}} \rewrites \ttt{OneProdMatchExp }A\expr{\ttt{$\lambda$ x $\rarr$ body}}\\
  &P\expr{\ttt{ $\rarr$ body}} \rewrites \ttt{NullaryMatch $M\expr{\ttt{body}}$}\\
  &\\
  &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} \rewrites \ttt{Lam (Name @a uniq) $M\expr{\ttt{body$[\ttt{x} \mapsto \ttt{Var uniq}]$}}$}\text{\;\;\;\;\;(where \ttt{uniq} is a globally unique identifier)}\\
  &\\
  &M\expr{\ttt{(x :: a)}} \rewrites
    \begin{cases}
      \ttt{externalize x} & \text{if}\; \nexists t, a\; \typeeq E\; t\\
      \ttt{x} & \text{if}\; \exists t, a\; \typeeq E\; t
    \end{cases}
\end{align*}
